---
title: "Homework 1"
author: "Abhishek Ajay (aa4266)"
date: "September 14, 2018"
output: html_document
---

```{r setup, include=FALSE}

#loading libraries

knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)

```
#Problem 1

Creating a data frame as instructed.

```{r problem_1_1}

#Creating data frame problem1_df

set.seed(1)

problem1_df = tibble( 
                        vec_numeric_col1 = runif(n = 10, min = 0, max = 5),
                        vec_logical_col2 = vec_numeric_col1 > 2,
                        vec_char_col3 = c("This", "is", "my", "first", "try", "with", "using", "R", "and", "RStudio"),
                        vec_factor_col4 = factor(c("Male", "Female", "Female", "Female", "Female", "Male", "Female", "Male", "Male", "Female"))
)

problem1_df

#Mean of each variable of our data frame, problem1_df

mean(problem1_df$vec_numeric_col1) #We get a mean value since the vector is of numeric data type

mean(problem1_df$vec_logical_col2) #We get a mean value since the logical variables are binomial with True = 1 and False = 0

mean(problem1_df$vec_char_col3) #Since the argument isn't numeric or logical, we get NA by coercion

mean(problem1_df$vec_factor_col4) #Since the argument isn't numeric or logical, we get NA by coercion

```

```{r problem_1_2}

#Explicit coercion of the logical, character and factor vectors of problem1_df to numeric type

problem1_df_vec_logical_col2_numeric = as.numeric(problem1_df$vec_logical_col2)

problem1_df_vec_char_col3_numeric = as.numeric(problem1_df$vec_char_col3)

problem1_df_vec_factor_col4_numeric = as.numeric(problem1_df$vec_factor_col4)

#The logical and factor vectors are transformed into numeric data type however, NAs are introduced by coercion into the character vector

```

```{r problem_1_3}

#Explicit sequential coercion of the character vectors of problem1_df to factor and then to numeric

problem1_df_vec_char_col3_factor_numeric = as.numeric(as.factor(problem1_df$vec_char_col3))

problem1_df_vec_char_col3_factor_numeric #The character column, when tranformed into factor and then numeric gives the index value of each entry as observed in the levels vector

#Explicit sequential coercion of the factor vectors of problem1_df to character and then to numeric

problem1_df_vec_factor_col4_char_numeric = as.numeric(as.character(problem1_df$vec_factor_col4))

problem1_df_vec_factor_col4_char_numeric #The factor column when tranformed into character and then numeric has NAs by coercion as its output since it has been converted to character first

```

In the above output we see that the character vector when converted to factor and then that to numeric gives a proper output of integers. However, the factor vector when converted to character and then numeric gives NA for each entry.

#Problem 2

```{r problem_2}

#Creating data frame problem2_df

set.seed(1)

problem2_df = tibble(
                    x = rnorm(1000),
                    y = rnorm(1000),
                    vec_logical_col3 = x + y > 0,
                    vec_numeric_col4 = as.numeric(vec_logical_col3),
                    vec_factor_col5 = as.factor(vec_logical_col3)
)

```

Short decription of the data frame created for problem 2: 

* The size of our data set is **`r dim(problem2_df)`**. 

* The mean of the first column, x, is **`r mean(problem2_df$x)`**. 

* The median of the first column, x, is **`r median(problem2_df$x)`**. 

<!-- calculating the proportion of cases where logical vector in problem2_df is true by dividing the length of the list of true responses in the logical vector by the total row count (i.e 1000) --> 

* The proportion of cases for which the logical vector is true: **`r length(problem2_df$vec_logical_col3[which(problem2_df$vec_logical_col3 == 'TRUE')])/nrow(problem2_df)`**.

```{r scatterplot_problem2_df}

#Creating scatter plots of data in problem2_df using the ggplot2 package and saving the first plot

ggplot(problem2_df, aes(x = x, y = y, color = vec_logical_col3)) + geom_point()

ggsave(filename = "C:/Users/abhis/OneDrive/Desktop/School/Semester 1/P8105/homework_1/p8105_hw1_aa4266/problem2_df_colour_logical_plot.png")

ggplot(problem2_df, aes(x = x, y = y, color = vec_numeric_col4)) + geom_point()

ggplot(problem2_df, aes(x = x, y = y, color = vec_factor_col5)) + geom_point()

```



