---
title: "Homework 1"
author: "Abhishek Ajay (aa4266)"
date: "September 14, 2018"
output: html_document
---

```{r setup, include=FALSE}

  knitr::opts_chunk$set(echo = TRUE)
  library(tidyverse)

```

#Problem 1

Creating a data frame as instructed.

```{r problem_1_1}

set.seed(1)

problem1_df = tibble( 
                        vec_numeric_col1 = runif(n = 10, min = 0, max = 5),
                        vec_logical_col2 = vec_numeric_col1 > 2,
                        vec_char_col3 = c("This", "is", "my", "first", "try", "with", "using", "R", "and", "RStudio"),
                        vec_factor_col4 = factor(c("Male", "Female", "Female", "Female", "Female", "Male", "Female", "Male", "Male", "Female"))
)
  
problem1_df

mean(problem1_df)

problem1_df_vec_logical_col2_numeric = as.numeric(problem1_df$vec_logical_col2)

problem1_df_vec_char_col3_numeric = as.numeric(problem1_df$vec_char_col3)

problem1_df_vec_factor_col4_numeric = as.numeric(problem1_df$vec_factor_col4)

```

```{r problem_1_2}

problem1_df_vec_char_col3_factor_numeric = as.numeric(as.factor(problem1_df$vec_char_col3))

problem1_df_vec_char_col3_factor_numeric

problem1_df_vec_factor_col4_char_numeric = as.numeric(as.character(problem1_df$vec_factor_col4))

problem1_df_vec_factor_col4_char_numeric

```

In the above output we see that the character vector when converted to factor and then that to numeric gives a proper output of integers. However, the factor vector when converted to character and then numeric gives NA for each entry.


```{r problem_2}

set.seed(1)

problem2_df = tibble(
                    x = rnorm(1000),
                    y = rnorm(1000),
                    vec_logical_col3 = x + y > 0,
                    vec_numeric_col4 = as.numeric(vec_logical_col3),
                    vec_factor_col5 = as.factor(vec_logical_col3)
)

```

Here we discuss about the various properties of our vector. *The size of our data set is `r dim(problem2_df)`.* The mean of the first column, x, is `r mean(problem2_df$x)` and its median is `r median(problem2_df$x)`. The proportion of cases for which the logical vector is true: `r length(problem2_df$vec_logical_col3[which(problem2_df$vec_logical_col3 == 'TRUE')])/nrow(problem2_df)`